Versinus: an Animated Visualization Method for Dynamic Scale-Free Networks\thanks{Supported by FAPESP, project 2017/05838-3}

### 2019-06-13 doc/paper.tex
primeira versao complea, falta lapidar

\begin{figure}[!h]\centering
\includegraphics[width=.9\textwidth]{sonif}
  \caption{Visual representation of the sonification implemented for the dynamic networks.
  $c_f$ and $f_0$ stands for \emph{center frequency} and \emph{fundamental frequency}, respectively.
  Further details are given in Section~\ref{sson}.
  }\label{fson}
\end{figure}

@@

\subsection{An experimental Visual Analytics interface}
This section exemplifies the employment of the Versinus method
within a more elaborate framework.
Let us consider an interactive visual interface, with which the user may parametrize Versinus and then navigate through the entailed dynamic network.

\begin{figure}[!h]\centering
\includegraphics[width=\textwidth]{esfNetVis___}
  \caption{
  }\label{esfnetvis}
\end{figure}

\begin{figure}[!h]\centering
\includegraphics[width=\textwidth]{esfNetVis2___}
  \caption{
  }\label{esfnetvis}
\end{figure}

\begin{figure}[!h]\centering
\includegraphics[width=\textwidth]{esfNetVis3}
  \caption{
  }\label{esfnetvis}
\end{figure}

\begin{figure}[!h]\centering
\includegraphics[width=\textwidth]{esfNetVis4}
  \caption{
  }\label{esfnetvis}
\end{figure}
One basic feature of the method is the constant representation of
nodes in each of the hubs, intermediary and peripheral sectors.
Nodes are positioned according to such classification regarding all
the activity of interest, and in each snapshot visual cues map the
classification within the snapshot.
Following the analytic methods used in~\cite{stab}, it may be crucial
to know how the relative sizes of the sectors vary along time, i.e. along
the sequence of snapshots.

>>>>

\begin{figure}[!h]\centering
\includegraphics[width=0.7\textwidth]{sonif}
  \caption{Visual representation of the sonification implemented for the dynamic networks.
  $c_f$ and $f_0$ stands for \emph{center frequency} and \emph{fundamental frequency}, respectively.
  Further details are given in Section~\ref{sson}.
  }\label{fson}
\end{figure}

@@

\subsection{An experimental visual analytics interface}
This section exemplifies the employment of the Versinus network visualization method
within a more elaborate framework.
Let us consider an interactive visual interface, with which the user may parametrize Versinus and then navigate through the entailed dynamic network.
A most fundamental Versinus parametrization is performed before the visual mapping of the network, and is illustrated in Figure~\ref{env1}.
Initial visualization is illustrated in Figure~\ref{env2}.
Figures~\ref{env3} and~\ref{env4} exemplifies Versinus compliant interactive data visualization analytical procedures: the observation of the fraction of nodes in each sector, characteristics of individual nodes and of snapshots.
The interactive interface is named $ESFNetVis$ (after ``Evolutive Scale Free Network Visualization''). At\footnote{Pivot repository for Versinus resources such as implementations, public instances of interactive interfaces, and videos: https://github.com/ttm/versinus} are directions for the code and a publicly accessible instance of $ESFNetVis$.

\begin{figure}[!h]\centering
\includegraphics[width=0.7\textwidth]{esfNetVis___}
  \caption{Visual elements for settings performed previous to the visual mapping of Versinus.
  The buttons on the top are for keeping, modifying and accessing settings.
  The Widgets in the gray rectangles are for selecting the network, selecting the first and last messages (transactions) to be considered,
  and the sectorialization method.
  The widgets in the white rectangle are $\sigma$, $\eta$, and $\omega$.
  These settings are compliant with the description available in Section~\ref{sec:des}.
  When the red button is clicked, the dynamic network is mapped to the canvas and additional elements, as described in Figure~\ref{env2}.
  }\label{env1}
\end{figure}

\begin{figure}[!h]\centering
\includegraphics[width=0.7\textwidth]{esfNetVis2___}
  \caption{The initial visualization of the dynamic network performed using $ESFNetVis$.
  In the canvas, the network is mapping using a 3D extension of Versinus, in which both segments oscillates vertically and in depth.
  The red portions of the links denotes the direction of the link (often performed using arrow heads).
  The text area starts with overall information of the collection of snapshots (e.g. each snapshot yields a mean degree, the set of snapshots yields a mean of such mean degree and a standard deviation of the mean degree.
  The timelines hold the fractions of nodes in each sector with orange, green and blue for hub, intermediary and peripheral nodes.
  The red bar is the playback cue, i.e. the snapshot being visualized.
  The analyst may play and pause with the play button,
  and rewind with the rewind button, both to the right.
  }\label{env2}
\end{figure}

\begin{figure}[!h]\centering
\includegraphics[width=0.7\textwidth]{esfNetVis3}
  \caption{The layout is slightly rotated in order to focus on the peripheral nodes segment. This image depicts one of the interaction possibilities among the canvas and the timelines: when the mouse is over a node, the line of the sector of the node in the snapshot is made black.
  (Notice the pointer in ``button-mode'', i.e. visualized as a hand.)
  Also, the text area in this image shows information about specific nodes, provided on-demand (the user hits the key ``i'' when the mouse is over a node for the interface to provide the information). 
  }\label{env3}
\end{figure}

\begin{figure}[!h]\centering
\includegraphics[width=0.7\textwidth]{esfNetVis4}
  \caption{
  The layout is rotated in order to focus on the intermediary segment of the sinusoid.
  The text area shows information about specific snapshots, given on-demand. The user requests the information by pressing "S" while the snapshot is being visualized.
  the nodes in the intermediary sectors are being highlighted on-demand by the user: notice the pointer which is over the intermediary sector timeline.
  }\label{env4}
\end{figure}
One basic feature of the method is the constant representation of
nodes in each of the hubs, intermediary and peripheral sectors.
Nodes are positioned according to such classification regarding all
the activity of interest, and in each snapshot visual cues map the
classification within the snapshot.
Following the analytic methods used in~\cite{stab}, it may be crucial
to know how the relative sizes of the sectors vary along time, i.e. along
the sequence of snapshots.
